Local to global N input modelling

### 2025-04-25 08_manuscript/_manuscript/08_manuscript.qmd
Updates

execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true

@@

@@

@@

@@

* Farms with mean temperature < 4$^\circ$ were excluded

@@

pop_params <- tar_read(pop_params, store = "../05_targets_farm")
species_params <- tar_read(species_params, store = "../05_targets_farm")
params <- c(pop_params, species_params)
pnames <- names(params)
sens_excr <- file.path("..", "data", "atlantic_salmon", "data_products", "excreted_parameter_sensitivity.parquet") %>%
  read_parquet() %>%
  mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
```

@@

|:-------:|:----:|:----:|----------------|-------------------------|

@@

| $\epsilon_{P}$ | `r unname(params['epsprot'])` | J g protein-1 | Energy content of protein | @baldan_r_2018 | 
| $\epsilon_{L}$ | `r unname(params['epslip'])` | J g lipid-1 | Energy content of lipid | @baldan_r_2018 |
| $\epsilon_{C}$ | `r unname(params['epscarb'])` | J g carbohydrate-1 | Energy content of carbohydrate | @baldan_r_2018 |
| $\epsilon_{O_2}$ | `r unname(params['epsO2'])` | J g O2-1 | Energy consumed by the respiration of 1 g of oxygen | @elliott_energy_1975 |
| $pk$ | `r unname(params['pk'])` | d-1 | Temperature coefficient for the fasting catabolism | @macnaughton_using_2019 |
| $k_0$ | `r unname(params['k0'])` | degree C-1 | Fasting catabolism at 0 Celsius degree | @macnaughton_using_2019 |

@@

| $T_{max}$ | `r unname(params['Tma'])` | C | Maximum lethal temperature | Juvenile Chinook salmon (19.8 $\pm$ 0.02 g, acclimation temp: 15$^{^\circ}$C, 19${\circ}$C) [@poletto_unusual_2017] |
| $T_{opt}$ | `r unname(params['Toa'])` | C | Optimal temperature | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$^{^\circ}$C [@jonsson_thermal_2001]. Alt value: 14°C [@handeland_effect_2008], *Salmo salar* initial size: 77.0 $\pm$ 14.6 g, temps: 6, 10, 14, 18${\circ}$C | 
| $T_{min}$ | `r unname(params['Taa'])` | C | Lowest feeding temperature | Fish essentially stop feeding when below 2$^{\circ}$C [@vadboncoeur_lowering_2023] | 

@@

| $a$ | `r unname(params['a'])` | J g tissue-1 | Energy content of fish tissue | Alt value: 17.02350 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |

@@

| $\overline{I_{max}}$ | `r unname(params['meanImax'])` | g g-1 fish day-1 | Ingestion rate mean | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\Delta{I_{max}}$ | `r unname(params['deltaImax'])` | g g-1 fish day-1 | Ingestion rate standard deviation | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $mort$ | `r unname(params['mortmyt'])` | d-1 | Natural mortality rate | @tvete_towards_2023 |

@@

sens_excr %>% 
  ggplot(aes(x = adj_param, y = sens, ymin = sens-sd, ymax = sens+sd)) +
  geom_col(alpha = 0.35, colour = "black") +
  geom_errorbar(width = 0.3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  scale_y_continuous(breaks = seq(-20,30,5)) +
  labs(x = "Adjusted parameter", y = "Sensitivity") +
  theme_classic() + theme(legend.position = "none", 
        strip.text = element_blank(), 
        aspect.ratio = 0.8,
        text = element_text(size = 12)) +
  coord_flip()
```

@@

farm_harvest_size <- tar_read(farm_harvest_size, store = "../04_targets_individual")

ggplot(farm_harvest_size, aes(x = weight/1000)) + 
  geom_histogram(colour = "black", fill = "salmon", alpha = 0.5) + 
  theme_classic() + 
  scale_x_continuous(breaks = seq(0, 8, 1), limits = c(1.25, 7.75)) + 
  labs(y = "Frequency", x = "Individual weight (kg)")
```

@@

>>>>

execute:
  eval: true
  echo: false
  warning: false
  message: false
  cache: true
  freeze: auto

@@

```{r global-variables}
#| include: false
#| eval: true

source(here("00_model_functions.R"))

indi_store <- here() %>% file.path("04_targets_individual")
farm_store <- here() %>% file.path("05_targets_farm")

species_params <- tar_read(species_params, store = indi_store)
pop_params <- tar_read(pop_params, store = farm_store)

farm_outputs <- here() %>% file.path("data", "atlantic_salmon", "data_products", "model_outputs_farm")
coho_outputs <- here() %>% file.path("data", "atlantic_salmon", "data_products", "model_outputs_cohort")
```

@@

[This is a note create with the `aside` class.]{.aside}

@@

```{r fig-farm-mean-temps}
#| fig-caption: Colours show farm site mean temperature

worldmap <- ne_countries(scale = "medium", returnclass = "sf")
farms_to_omit <- here() %>% 
  file.path("data", "_general_data", "farm_locations", "atlantic_salmon_farms_to_omit.qs") %>% 
  qread()

farms_w_temp_df <- here() %>% 
  file.path("data", "_general_data", "farm_locations", "atlantic_salmon_locations_w_temps.qs") %>% 
  qread() %>% 
  filter(!farm_id %in% farms_to_omit) %>% 
  mutate(country = as.factor(country),
         species_group = as.factor(species_group),
         data_type = as.factor(data_type),
         data_type_2 = as.factor(data_type_2),
         model_name = as.factor(model_name),
         F_CODE = as.factor(F_CODE),
         day = str_split_i(day, "_", 2) %>% as.integer())

farm_temp_means <- farms_w_temp_df %>% 
  select(iso3c, country, farm_id, temp_c)
farm_temp_means_1 <- farms_w_temp_df %>% 
  filter(day == 1) %>% 
  select(geometry)

farm_temp_means <- farm_temp_means %>% 
  group_by(iso3c, country, farm_id) %>% 
  reframe(temp_c = meanna(temp_c))

farm_temp_means <- cbind(farm_temp_means, farm_temp_means_1)

ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = temp_c), size = 1.5) +
  coord_sf() +
  theme_classic() 
```

```{r fig-farm_temp_means-2}
farm_temp_means %>%
  mutate(
    coords = st_coordinates(geometry),
    longitude = coords[,1],
    latitude = coords[,2]
  ) %>%
  select(-coords) %>% 
  ggplot(aes(y = temp_c, x = abs(latitude), colour = country)) +
  geom_point() +
  labs(x = expression("Absolute latitude ("*degree*")"), y = expression("Temperature ("*degree*"C)")) +
  theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

* Farms with a mean temperature of $\leq 6^\circ$C were excluded.
```{r farms omitted}
farms_to_omit <- here() %>% file.path("data/_general_data/farm_locations/atlantic_salmon_farms_to_omit.qs") %>% 
  qread()
```

[Need to put the cohort concept (and harvest timing) in here.]{.aside}

```{r fig-cohorts}
#| fig-cap: Example of cohorts

fnms <- coho_outputs %>% 
  list.files(full.names = T) %>% 
  str_subset("biomass")

df <- fnms[1] %>% read_parquet() %>% 
  filter(feed == "reference") %>% 
  mutate(mean = set_units(mean, "g") %>% set_units("t"),
         sd = set_units(sd, "g") %>% set_units("t"))

ggplot(df, aes(x = yday, y = mean, ymin = mean-sd, ymax = mean+sd, colour = as.factor(cohort))) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(linetype = "dotted", alpha = 0) +
  scale_x_continuous(breaks = seq(0, 2000, 200)) +
  labs(y = "Farm biomass", x = "Day of the year") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 0.75)

yrs <- max(df$yday)/365
harvests <- df %>% 
  group_by(cohort) %>% 
  reframe(mean = maxna(mean),
          sd = maxna(sd))

sum(harvests$mean)/yrs
mean(harvests$sd)/yrs
```

### Feeding rate

```{r fig-functional-response-to-temperature}
#| fig-cap: Temperature changes in potential ingestion (solid line) and amount of food provided (dashed line), relative to maximum potential ingestion.

df <- data.frame(water_temp = seq(-5, 26, 0.25))
df$feeding <- sapply(FUN = feeding_rate, X = df$water_temp, species_params = species_params)
df <- df %>% 
mutate(ing_pot = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * feeding,
       food_prov = NA)
for (i in 1:nrow(df)) {
  df$food_prov[i] <- food_prov_rate(pop_params = pop_params, rel_feeding = df$feeding[i], ing_pot = df$ing_pot[i], 
                                    ing_pot_10 = pop_params['meanImax'] * (pop_params['meanW']^species_params['m']) * 0.1)
}
df$ing_pot <- df$ing_pot/maxna(df$ing_pot)
df$food_prov <- df$food_prov/maxna(df$ing_pot)

df %>% 
  ggplot(aes(x = water_temp, y = ing_pot)) +
  geom_line(linewidth = 0.75) +
  geom_line(aes(y = food_prov), linewidth = 0.75, linetype = "dotted") +
  scale_y_continuous(breaks = seq(0,1.5,0.25)) +
  scale_x_continuous(breaks = seq(-5,30,5), limits = c(-5,26)) +
  labs(x = expression("Water temperature ("*degree*"C)"), y = "Relative food ingested/provided") +
  theme_classic() +
  theme(text = element_text(colour = "black", family = "serif", size = 12))
```

@@

params <- c(pop_params, species_params)
pnames <- names(params)
sens_excr <- here() %>% file.path("data", "atlantic_salmon", "data_products", "excreted_parameter_sensitivity.parquet") %>%
  read_parquet() %>%
  mutate(adj_param = factor(adj_param, levels = rev(adj_param)))
```

@@

|:-------:|:----:|:----:|:----------------|:-------------------------|

@@

| $\epsilon_{P}$ | `r unname(params['epsprot'])` | J g protein$^{-1}$ | Energy content of protein | @baldan_r_2018 | 
| $\epsilon_{L}$ | `r unname(params['epslip'])` | J g lipid$^{-1}$ | Energy content of lipid | @baldan_r_2018 |
| $\epsilon_{C}$ | `r unname(params['epscarb'])` | J g carbohydrate$^{-1}$ | Energy content of carbohydrate | @baldan_r_2018 |
| $\epsilon_{O_2}$ | `r unname(params['epsO2'])` | J g O2$^{-1}$ | Energy consumed by the respiration of 1 g of oxygen | @elliott_energy_1975 |
| $pk$ | `r unname(params['pk'])` | d$^{-1}$ | Temperature coefficient for the fasting catabolism | @macnaughton_using_2019 |
| $k_0$ | `r unname(params['k0'])` | - | Fasting catabolism at 0 $^{^\circ}$C | @macnaughton_using_2019 |

@@

| $T_{max}$ | `r unname(params['Tma'])` | $^{^\circ}$C | Maximum lethal temperature | Juvenile *Oncorhynchus tshawytscha* (19.8 $\pm$ 0.02 g, acclimation temp: 15$^{^\circ}$C, 19${\circ}$C) [@poletto_unusual_2017] |
| $T_{opt}$ | `r unname(params['Toa'])` | $^{^\circ}$C | Optimal temperature | Fish size range: 2.25 $\pm$ 0.49 – 11.5 $\pm$ 5.29 g, temperature ranges: 5–24$^{^\circ}$C [@jonsson_thermal_2001]. Alt value: 14°C [@handeland_effect_2008], *Salmo salar* initial size: 77.0 $\pm$ 14.6 g, temps: 6, 10, 14, 18${\circ}$C | 
| $T_{min}$ | `r unname(params['Taa'])` | $^{^\circ}$C | Lowest feeding temperature | Fish essentially stop feeding when below 2$^{\circ}$C [@vadboncoeur_lowering_2023] | 

@@

| $a$ | `r unname(params['a'])` | J g tissue$^{-1}$ | Energy content of fish tissue | Alt value: 17.02350 (nls model fitted to female Atlantic salmon data extracted and transformed from @jonsson_energy_2003 [@van_tien_development_2016] |

@@

| $\overline{I_{max}}$ | `r unname(params['meanImax'])` | g g$^{-1}$ fish day$^{-1}$ | Ingestion rate mean | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $\Delta{I_{max}}$ | `r unname(params['deltaImax'])` | g g$^{-1}$ fish day$^{-1}$ | Ingestion rate standard deviation | [FAO](https://www.fao.org/fileadmin/user_upload/affris/docs/tabl10.pdf) |
| $mort$ | `r unname(params['mortmyt'])` | d$^{-1}$ | Natural mortality rate | @tvete_towards_2023 |

@@

sens_excr %>% 
  ggplot(aes(x = adj_param, y = sens, ymin = sens-sd, ymax = sens+sd)) +
  geom_col(colour = "black", fill = "salmon", alpha = 0.5) +
  geom_errorbar(width = 0.3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  scale_y_continuous(breaks = seq(-20,30,5)) +
  labs(x = "Adjusted parameter", y = "Sensitivity") +
  theme_classic() + theme(legend.position = "none", 
        strip.text = element_blank(), 
        aspect.ratio = 0.8,
        text = element_text(size = 12)) +
  coord_flip()
```

@@

pdata <- here() %>% file.path("data", "atlantic_salmon", "data_products", "harvest_size.parquet") %>% read_parquet()
pdata %>% 
  mutate(weight = weight %>% set_units("g") %>% set_units("kg")) %>% 
  select(c(farm_ID, weight)) %>% 
  ggplot(aes(x = weight)) +
  geom_histogram(binwidth = 0.25, colour = "black", fill = "salmon", alpha = 0.5) +
  scale_y_continuous(limits = c(0,500)) +
  labs(y = "Frequency", x = "Harvest weight") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 0.75)
```

@@

# Results

* Main point is the difference between feeds. 
* This should be expressed in t/t salmon, so farm biomass is also important.
* Also interesting to see if difference between feeds varies geographically (correlates with mean/median/max temperature?)

# Biomass produced

```{r biomass produced}
fnms <- farm_outputs %>% 
  list.files(full.names = T) %>% 
  str_subset("biomass")

dfs <- list()
for (fid in seq_along(fnms)) {
  dfs[[fid]] <- fnms[fid] %>% 
    read_parquet() %>% 
    filter(days == max(days)) %>% 
    mutate(mean = mean %>% set_units("g") %>% set_units("t"),
           sd = sd %>% set_units("g") %>% set_units("t"))
}
df <- bind_rows(dfs)
df_mean <- df %>% 
  select(-c(days, sd)) %>% 
  pivot_wider(names_from = feed, values_from = mean) %>% 
  mutate(future_diff = (future - reference)/reference,
         past_diff = (past - reference)/reference) %>% 
  select(-c(reference, past, future)) %>% 
  pivot_longer(names_to = "feed", values_to = "mean", cols = c(future_diff, past_diff))
df_sd <- df %>% 
  select(-c(days, mean)) %>% 
  pivot_wider(names_from = feed, values_from = sd) %>% 
  mutate(future_diff = (future - reference)/reference,
         past_diff = (past - reference)/reference) %>% 
  select(-c(reference, past, future)) %>% 
  pivot_longer(names_to = "feed", values_to = "sd", cols = c(future_diff, past_diff))
df <- merge(df_mean, df_sd, by = c("farm_ID", "feed"))
```

# Locations
